\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-07-17 introduction.tex
tweaks

There is a growing trend among semantics researchers to
use proof assistants to mechanize meta\-theories.
Consequently, it has become ever more important that proof assistants
make it easy to write reusable, extensible code and proofs.

However, the programmer runs into an old problem,
the Expression Problem (EP) \cite{wadler-ep},
that has deep roots in language design.
%
The EP is a programming challenge
epitomizing the difficulty of writing type-safe, extensible code.
To define an expression language that can be reused for future extension,
the programmer faces a fundamental tension \cite{reynolds1975} between
adding new constructors of a data type (e.g., new abstract syntax) and
adding new functions over the data type (e.g., new compiler passes).

@@

In contrast, proof assistants offer few linguistic solutions to
programmers who want to mechanize their metatheories.
Yet the challenge of writing extensible, type-safe code (and proofs!) is
as real.
As an example, consider \cref{fig:STLC-example}, the simply typed
lambda calculus (STLC) mechanized in Coq.
A programming challenge is to define distinct extensions of this STLC
formalization to support distinct features,
and then selectively compose these extensions to form new STLC variants.

@@

In particular, when a language mechanism applies inheritance to
recursive, record-like definitions containing related types and terms,
it is said to support \emph{family polymorphism}~\cite{ernst2001family}:
code is polymorphic to a family it is nested within.

@@

We contribute a language design that integrates family polymorphism into
a proof assistant.
Because code and proofs are polymorphic to a family it is nested
within, they can be inherited and reused by a derived family.
Extensible metatheory mechanization is thus made possible.

Integrating family polymorphism into a dependent type theory poses
significant technical challenges, however.
\TODO

>>>>

There is a growing trend among programming languages researchers
to use proof assistants to mechanize meta\-theories.
%
However, the programmer runs into an old problem,
the Expression Problem (EP) \cite{wadler-ep},
in the new setting of proof engineering.

The EP is a programming challenge that
epitomizes the difficulty of writing type-safe, extensible code.
To define an expression language that can be reused for future extension,
the programmer faces a fundamental tension \cite{reynolds1975} between
adding new constructors of a data type (e.g., new abstract syntax) and
adding new functions over the data type (e.g., new compiler passes).

@@

In contrast, proof assistants offer few linguistic solutions that
address the EP.
Yet the challenge of writing extensible, type-safe code (and proofs!) is
as real, especially for metatheory mechanization.
As an example, consider \cref{fig:STLC-example}, the simply typed
lambda calculus (STLC) mechanized in Coq.
A programming challenge there is to define distinct extensions of this
STLC formalization to support distinct features,
and then selectively compose these extensions to form new STLC variants.

@@

In particular, when a language mechanism enables inheritance over
recursive, record-like definitions containing related types and terms,
it is said to support \emph{family polymorphism}~\cite{ernst2001family}:
code is polymorphic to a family it is nested within.

@@

We contribute a language design that integrates family polymorphism into
a proof assistant.
Because code and proofs are polymorphic to a family it is nested
within, they can be inherited and reused by a derived family.
%
Family polymorphism enables extensible metatheory mechanization.
As \cref{fig:?} shows,
different extensions to STLC all inherit from the base STLC family:
they reuse proofs of metatheories including the progress and
preservation theorems, only adding new cases to the induction proofs as
needed by the extension.

Integrating family polymorphism into a dependent type theory with
inductive types poses significant technical challenges, however.
\TODO
